---
title: 'NYT Bestsellers: 03 Tops and Ranks'
author: "A.H. Wilhelm"
date: "8/22/2018"
output:
  html_document: default
  pdf_document: default
---


###I created a couple of data tables to see tops.

##Loading the data.

```{r loading, warning=F, message=F}
library(readr)
library(ggplot2)
library(forcats)
library(dplyr)
library(ggthemes) 
library(DT)
library(tidyr)
library(stringr)
#create dataframes from new data files
df_nytinfo_rank <- read_csv("data/output/nytinfo_rank.csv")
df_nytinfo_weeks <- read_csv("data/output/nytinfo_weeks.csv")
```

##Create data tables

Create a data table of all titles. (Sorting this will show some confusion with "The Help", which is disappointing! I believe it is caused because it is listed under two different ISBNs, but I'm too tired to figure out where that occurred...)
```{r create_datatable1, warning=F, message=F}

#subset for DT
nytsort <- select(df_nytinfo_weeks, Title_NY, Category, Author_NY, weeks, type)
datatable(nytsort)

```




#The 5 fiction books that stayed on the list the longest were:
```{r create_topFtable, warning=F, message=F}
#top 5 fiction
nyttopsf <- nytsort %>% 
  filter(type=="fiction") %>% 
  arrange(desc(weeks)) %>% 
  head(5) 
  
datatable(nyttopsf)

```
Wow, "The Help" and "A Dance With Dragons" were on the list for more than two years. TWO YEARS!!! That's a long freaking time! And seeing as how this is the first time I've ever heard of "A Dance With Dragons", I'm starting to wonder if I even know all that much about books. 





#The top 5 nonfiction books were:
```{r create_topNFtable, warning=F, message=F}
#top 5 nonfiction
nyttopsnf <- nytsort %>% 
  filter(type=="nonfiction") %>% 
  arrange(desc(weeks)) %>% 
  head(5) 

  datatable(nyttopsnf)

```
I'd forgotten all about "Unbroken", but wow what a book that was. 
