---
title: 'NYT Bestsellers: 01 Data Cleanup'
author: "A.H. Wilhelm"
date: "8/22/2018"
output:
  html_document: default
  pdf_document: default
---


###This set of charts is based on data I found online from another group's work (which was really awesome, by the way).

I worked with many packages to wrangle the data. Among other things, I filled in holes for author_gender, and joined data to create new values like totals and averages. In the end I also filtered out "multi-author" books so I could do a clean gender compare. 

This sounds so simple! But as a newbie it took me a mill lines of code.

##Loading packages and data.

```{r loading, warning=F, message=F}
library(readr)
library(ggplot2)
library(forcats)
library(dplyr)
library(ggthemes) 
library(DT)
library(tidyr) 
library(stringr)

#create dataframes from original data files
df_nytfic_isbntall<-read.csv("data/originals/nytb2008-2016Fiction.csv")
df_nytnonfic_isbntall<-read.csv("data/originals/nytb2008-2016NonFiction.csv")
df_nyt_info<-read.csv("data/originals/isbnToInfo.csv")

```

##Cleaning data
I cleaned up the data a bit to fill some holes I noticed once I began to visualize it. Here I am filling in some missing genders and correcting some fiction/nonfiction things. 

```{r cleaning_data, warning=F, message=F}

df_nyt_info2 <- df_nyt_info %>% 
  mutate(author_gender=case_when(
    Author_NY %in% c("James Patterson and Maxine Paetro", "the editors of Life magazine", 
                     "W R Bolen and the creators of TotalFratMovecom", "Rorke Denver and Ellis Henican",
                     "the staff of The Onion", "Tim Howard with Ali Benjamin", "the editors of Life Magazine",
                     "Howard Schultz and Rajiv Chandrasekaran" )~"multi",
    Author_NY %in% c("Kresley Cole", "J D Robb", "J R Ward", "Kathryn Stockett", "Danielle Steel", "Jhumpa Lahiri",
                     "Lisa Lutz", "M L Stedman", "Téa Obreht", "J Courtney Sullivan", "JR Ward",
                     "Karen Kingsbury")~"female",
    Author_NY %in% c("Bernard Cornwell","Bernard Cornwell","D T Max", "S C Gwynne", 
                     "A J Baime", "M O Walsh","Glennon Doyle Melton", "T Boone Pickens", "Rinker Buck", "David Weber", "T C Boyle",
                     "T Coraghessan Boyle", "Jon Ronson", "Nassir Ghaemi")~"male",
    TRUE ~ as.character(author_gender)
    )) %>% 
  mutate(Category=case_when(
    Author_NY=="Karen Kingsbury" ~"Religious Fiction", 
    Author_NY=="David Sedaris" & Title_NY=="HOLIDAYS ON ICE"~"Nonfiction General",
    TRUE ~ as.character(Category)
    ))

```

I wanted to convert the "wide"" isbn file to "tall" so that I could do other things with the data like tabulate number of weeks a book stayed on the list, and also join it with the information about the books.
```{r spreading_data, warning=F, message=F}
#fic and nonfic: make isbn/weeks tall 
df_nytfic_isbntall2<- df_nytfic_isbntall%>% 
  gather(key="number", value="isbn",2:21)
df_nytnonfic_isbntall2<- df_nytnonfic_isbntall%>% 
  gather(key="number", value="isbn",2:21)

#fic and nonfic: total number of weeks per isbn/title
df_nytfic_isbnweeks<- df_nytfic_isbntall2 %>% 
  group_by(isbn) %>% 
  summarize(weeks=n()) %>% 
  mutate(type="fiction")

df_nytnonfic_isbnweeks<- df_nytnonfic_isbntall2 %>% 
  group_by(isbn) %>% 
  summarize(weeks=n()) %>% 
  mutate(type="nonfiction")

#join weeks
df_both_isbnweeks <- union(df_nytfic_isbnweeks, df_nytnonfic_isbnweeks)

#Replace X# ranks with numbers, to have rank available for line chart y axis.
df_nytfic_isbntall2<-df_nytfic_isbntall2 %>% 
  mutate(number = as.character(number),
         number = if_else(number =="X1", "01", number ),
         number = if_else(number =="X2", "02", number ),
         number = if_else(number =="X3", "03", number ),
         number = if_else(number =="X4", "04", number ),
         number = if_else(number =="X5", "05", number ),
         number = if_else(number =="X6", "06", number ),
         number = if_else(number =="X7", "07", number ),
         number = if_else(number =="X8", "08", number ),
         number = if_else(number =="X9", "09", number ),
         number = if_else(number =="X10", "10", number ),
         number = if_else(number =="X11", "11", number ),
         number = if_else(number =="X12", "12", number ),
         number = if_else(number =="X13", "13", number ),
         number = if_else(number =="X14", "14", number ),
         number = if_else(number =="X15", "15", number ),
         number = if_else(number =="X16", "16", number ),
         number = if_else(number =="X17", "17", number ),
         number = if_else(number =="X18", "18", number ),
         number = if_else(number =="X19", "19", number ),
         number = if_else(number =="X20", "20", number ))

df_nytnonfic_isbntall2<-df_nytnonfic_isbntall2 %>% 
  mutate(number = as.character(number),
         number = if_else(number =="X1", "01", number ),
         number = if_else(number =="X2", "02", number ),
         number = if_else(number =="X3", "03", number ),
         number = if_else(number =="X4", "04", number ),
         number = if_else(number =="X5", "05", number ),
         number = if_else(number =="X6", "06", number ),
         number = if_else(number =="X7", "07", number ),
         number = if_else(number =="X8", "08", number ),
         number = if_else(number =="X9", "09", number ),
         number = if_else(number =="X10", "10", number ),
         number = if_else(number =="X11", "11", number ),
         number = if_else(number =="X12", "12", number ),
         number = if_else(number =="X13", "13", number ),
         number = if_else(number =="X14", "14", number ),
         number = if_else(number =="X15", "15", number ),
         number = if_else(number =="X16", "16", number ),
         number = if_else(number =="X17", "17", number ),
         number = if_else(number =="X18", "18", number ),
         number = if_else(number =="X19", "19", number ),
         number = if_else(number =="X20", "20", number ))

#combine fic_ and nonfic_ isbntall frames into one data frame
df_both_isbntall <- union(df_nytfic_isbntall2, df_nytnonfic_isbntall2)


```

And now that I have the datasets in the shape I want, I am writing them out.
```{r writing_files, warning=F, message=F}
#****Create and write out data sets****#

#join df_both_isbnweeks & df_nytinfo, WRITE CSV
df_nytinfo_weeks<-left_join(df_nyt_info2, df_both_isbnweeks, by="isbn")
write_csv(df_nytinfo_weeks, "nytinfo_weeks.csv")

#join df_both_isbntall & df_nytinfo_weeks, WRITE CSV
df_nytinfo_rank<-left_join(df_nytinfo_weeks, df_both_isbntall, by="isbn")
write_csv(df_nytinfo_rank, "nytinfo_rank.csv")

```

